---
title: "Census Data Clustering ZIPs"
author: "JR Morrish"
date: "8/6/2020"
---

```{r packages, echo=FALSE, include=FALSE}
# install.packages("tinytex")
# library(tinytex)
# tinytex::install_tinytex()
library(ape)        # Hierarchical clustering
library(cluster)    # Clustering algorithms
library(clustertend)# Assess cluster tendency
library(clv)        # Evaluate stability of clustering solution
library(clValid)    # Evaluate stability of clustering solution
library(dbscan)     # Clustering algorithms
library(dendextend) # Modify graphs
library(dplyr)      # Data manipulation
library(factoextra) # Clustering algorithms & visualization
library(fpc)        # Evaluate stability of clustering solution
library(ggplot2)    # Create graphs
library(gridExtra)  # Multiple cluster plots in grid
#library(multiClust) # Data manipulation
library(mvtnorm)    # For other packages
library(mvoutlier)  # Identify multivariate outliers
library(NbClust)    # Determine number of clusters
library(outliers)   # Test for univariate outliers
library(psych)      # Descriptive statistics
library(pvclust)    # Compute cluster p-values
library(readxl)     # Read in excel files
library(Rtsne)      # Clustering mixed data
library(tidyverse)  # Data manipulation
library(openxlsx)   # Write excel files
# library(ggstatsplot)
# tinytex:::is_tinytex()
```

census <- readxl::read_xlsx("combined_data.xlsx", sheet = "Census")

census <- na.omit(census)
ef <- na.omit(census)

census$ZIP_Pop_Density <- as.numeric(census$ZIP_Pop_Density)
census$ZIP_med_income <- as.numeric(census$ZIP_med_income)
census$ZIP_med_age <- as.numeric(census$ZIP_med_age)
census$ZIP_P_Black <- as.numeric(census$ZIP_P_Black)
census$ZIP_P_Asian <- as.numeric(census$ZIP_P_Asian)
census$ZIP_P_Hispanic <- as.numeric(census$ZIP_P_Hispanic)

#### Transformed Data
```{r log transform data, echo=FALSE, include=FALSE}
a <- min(census$ZIP_Pop_Density[census$ZIP_Pop_Density>0])-.0001
g <- min(census$ZIP_P_Black[census$ZIP_P_Black>0])-.0001
h <- min(census$ZIP_P_Asian[census$ZIP_P_Asian>0])-.0001
i <- min(census$ZIP_P_Hispanic[census$ZIP_P_Hispanic>0])-.0001

census %>%
  mutate( 
    pd.log = log(ifelse(ZIP_Pop_Density==0,a,ZIP_Pop_Density)),
    pBl.log = log(ifelse(ZIP_P_Black==0,g,ZIP_P_Black)),
    pAn.log = log(ifelse(ZIP_P_Asian==0,h,ZIP_P_Asian)),
    pHL.log = log(ifelse(ZIP_P_Hispanic==0,i,ZIP_P_Hispanic))
  ) -> census
```

```{r normalize data, echo=FALSE, include=FALSE}
sccensus <- scale(census)

sum(is.na(sccensus))
# list rows of data that have missing values
sccensus[!complete.cases(sccensus),]
```

```{r cluster data}
sccensus <- census %>% 
  dplyr::select(2:3, 7:10) %>% 
  scale()

set.seed(4323)

sccensus %>%
  dist() %>%
  hclust() -> census.hclust

# R16 <- ef %>%
#   mutate(cluster = cutree(census.hclust,16))

# as.phylo(census.hclust) -> hcc
# plot(hcc,type="phylogram",cex=.5,label.offset=.1)
# plot(hcc,type="unrooted", cex=0.2,label.offset=.1)
fviz_dend(census.hclust, k = 16, rect = TRUE, rect_fill = TRUE, lwd = 0.5, cex = 0.5) # pop_dens, med_age, %s, med_inc
# fviz_dend(census.hclust, k = 16, rect = TRUE, rect_fill = TRUE, xlim = c(0,100), ylim = c(0,8))
# pdf("30.pdf")
``` 

```{r hierachical cuts}
# R16 <- ef %>%
#   mutate(cluster = cutree(census.hclust,3))
R16 <- ef %>%
 mutate(cluster = cutree(census.hclust3,16))

# R30 <- ef %>%
#  mutate(cluster = cutree(census.hclust3,30))
```


```{r export, echo=FALSE, include=FALSE, eval=FALSE}
metrics <- list(
  min  = ~min(.x, na.rm = TRUE), 
  max  = ~max(.x, na.rm = TRUE),
  mean = ~mean(.x, na.rm = TRUE),
  sd   = ~sd(.x, na.rm = TRUE)
)

R16stats <- R16 %>%
  group_by(cluster) %>%
  summarise(
    count = n(),
    pctTot = count/nrow(R16),
    across(9:26, metrics)) %>% 
  select(cluster, count, pctTot,
         Pop_Density_mean,
         med_income_mean,
         med_age_mean,
         P_Black_mean,
         P_Asian_mean,
         P_Hispanic_mean)

v3 <- createWorkbook()

addWorksheet(v3,"Clusters")
addWorksheet(v3,"STATS")

writeData(v3, 1, R16)
writeData(v3, 2, R16stats)

saveWorkbook(v3, 
             file="C:/Users/jr_morrish/Documents/Projects/Cluster Analysis/Customers/output/Census Clusters.xlsx", overwrite = TRUE)
```
